---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Set working directory and library packages

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
setwd("D:/Coursera")
Sys.setlocale("LC_TIME", "English")
library(RColorBrewer)
library(ggplot2)
library(knitr)
library(dplyr)
```

Load the data to a local file and unzip the data

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
fileURL<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileURL,destfile = "./Activity.zip")
listZip<-unzip("./Activity.zip")
```

Read the data

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
activity<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day and make a histogram of the total number of steps taken each day

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
daysum<-summarize(group_by(activity,date),sum=sum(steps,na.rm = T))
gsum<-ggplot(data=daysum,aes(sum))
gsum+geom_histogram(bins=30,fill=colorRampPalette(brewer.pal(3,"Dark2"))(1))+ggtitle(label = "Total Number of Steps Taken Each Day")
```

Calculate and report the mean and median of the total number of steps taken per day

```{r}
stepmean<-round(mean(daysum$sum),2)
stepmean
stepmedian<-median(daysum$sum)
stepmedian
```

The mean of the total number of steps taken per day was `r stepmean` and the median was `r stepmedian`

## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
interaver<-summarize(group_by(activity,interval),average=mean(steps,na.rm = T))
gaver<-ggplot(data=interaver,aes(x=interval, y=average))
gaver+geom_line(linetype=1,color=colorRampPalette(brewer.pal(9,"RdGy"))(1))+ggtitle(label = "Average Number of Steps Taken Across All Days")+coord_cartesian(xlim=c(0,2400))
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
maxinterval<-filter(interaver,average==max(interaver$average))[,1]
maxinterval
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
sum(as.integer(is.na(activity$steps)))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in with the mean for that 5-minute interval

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
actfill<-activity
for(i in seq_len(nrow(actfill))){
  if(is.na(actfill$steps[i])){
    actfill$steps[i]<-as.integer(filter(interaver, interaver$interval==actfill$interval[i])[,2])
  }
}
```

Make a histogram of the total number of steps taken each day

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
daysumfill<-summarize(group_by(actfill,date),sum=sum(steps,na.rm = T))
gsumfill<-ggplot(data=daysumfill,aes(sum))
gsumfill+geom_histogram(bins=30,fill=colorRampPalette(brewer.pal(9,"PRGn"))(1))+ggtitle(label = "Total Number of Steps Taken Each Day after Filling in Missing Values")
```

Calculate and report the mean and median total number of steps taken per day

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
stepmeanfill<-round(mean(daysumfill$sum),2)
stepmeanfill
stepmedianfill<-median(daysumfill$sum)
stepmedianfill
```

After inputing missing data on the estimates of the total daily number of steps, both mean and median number of steps taken per day was increased. Mean number increased from `r stepmean` to `r stepmeanfill` and median number increased from `r stepmedian` to `r stepmedianfill`

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels ¨C ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
actfillwe<-mutate(actfill,weekday=as.factor(ifelse(weekdays(as.Date(date))=="Sunday","weekend",ifelse(weekdays(as.Date(date))=="Saturday","weekend","weekday"))))
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
interaverfill<-summarize(group_by(actfillwe,weekday,interval),average=mean(steps,na.rm = T))
gaverfill<-ggplot(data=interaverfill,aes(x=interval, y=average))
gaverfill+facet_wrap(~weekday, nrow=2)+geom_line(linetype=1,color=colorRampPalette(brewer.pal(6,"PiYG"))(1))+ggtitle(label = "Average Number of Steps Taken Across Weekday and Weekend")+labs(y="Average Number of Steps")
```
